---
title: "Chapter 5"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r include=FALSE}
# Setting Knit options for the document
knitr::opts_chunk$set( message=FALSE, warning=FALSE)
```
\newcommand{\indep}{\perp \!\!\! \perp}


# 5.1: Supurious Correlations
## We observe a puzzle: Marriage rates and divorce rates are positively correlated
```{r}
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce

d$W <- d$WaffleHouses
d$A <- scale(d$MedianAgeMarriage)
d$D <- scale(d$Divorce)
d$M <- scale(d$Marriage)
```

Let's fit a bayesian model to the three variables -- Waffle Houses, Age at marriage and Marriage Rate.

```{r}
waffle_divorce <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + bW*W,
    bW ~ dnorm(0, 0.2),
    a ~ dnorm(0, 0.2),
    sigma ~ dexp(1)
  ),
  data = d
)
age_divorce <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + bA*A,
    bA ~ dnorm(0, 0.2),
    a ~ dnorm(0, 0.2),
    sigma ~ dexp(1)
  ),
  data = d
)

marriage_divorce <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + bM*M,
    bM ~ dnorm(0, 0.2),
    a ~ dnorm(0, 0.2),
    sigma ~ dexp(1)
  ),
  data = d
)
```

```{r}
waffle_divorce_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(waffle_divorce) , Sigma=vcov(waffle_divorce))
waffle_seq <- seq(from=min(d$W)-1, to = max(d$W) + 1, length.out = 30)
input_matrix <- rbind(rep(1,length(waffle_seq)), waffle_seq)
mean_divorce_sim_m5_1 <- waffle_divorce_posterior[,c("a","bW")] %*% input_matrix
mean_divorce_PI_m5_1 <- apply(mean_divorce_sim_m5_1, 2, rethinking::PI, prob=0.89)

age_divorce_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(age_divorce) , Sigma=vcov(age_divorce))
age_seq <- seq(from=min(d$A)-1, to = max(d$A) + 1, length.out = 30)
input_matrix <- rbind(rep(1,length(age_seq)), age_seq)
mean_divorce_sim_m5_2 <- age_divorce_posterior[,c("a","bA")] %*% input_matrix
mean_divorce_PI_m5_2 <- apply(mean_divorce_sim_m5_2, 2, rethinking::PI, prob=0.89)

marriage_divorce_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(marriage_divorce) , Sigma=vcov(marriage_divorce))
marriage_seq <- seq(from=min(d$M)-1, to = max(d$M) + 1, length.out = 30)
input_matrix <- rbind(rep(1,length(marriage_seq)), marriage_seq)
mean_divorce_sim_m5_3 <- marriage_divorce_posterior[,c("a","bM")] %*% input_matrix
mean_divorce_PI_m5_3 <- apply(mean_divorce_sim_m5_3, 2, rethinking::PI, prob=0.89)

# Plot the graphs
par(mfcol=c(1,3))

plot(d$WaffleHouses, d$D, main="Waffle Houses and Divorce Rate")
abline(a=mean(waffle_divorce_posterior[,"a"]), b=mean(waffle_divorce_posterior[,"bW"]), lty=2)
shade(mean_divorce_PI_m5_1, waffle_seq)

plot(d$A, d$D, main="Age at marriage and Divorce Rate")
abline(a=mean(age_divorce_posterior[,"a"]), b=mean(age_divorce_posterior[,"bA"]), lty=2)
shade(mean_divorce_PI_m5_2, age_seq)

plot(d$M, d$D, main="Marriage Rate and Divorce Rate")
abline(a=mean(marriage_divorce_posterior[,"a"]), b=mean(marriage_divorce_posterior[,"bM"]), lty=2)
shade(mean_divorce_PI_m5_3, marriage_seq)
```


## Introduce some tools to solve the puzzle: DAGs
A DAG is a way of describing qualitative causal relationships among variables. It isn’t as detailed as a full model description, but it contains information that a purely statistical model does not. Unlike a statistical model, a DAG, if it is correct, will tell you the consequences of intervening to change a variable.

```{r}
library(dagitty)
dag5.1 <- dagitty( "dag {
    A -> D
    A -> M
    M -> D
}")
coordinates(dag5.1) <- list( x=c(A=0,D=1,M=2) , y=c(A=0,D=1,M=0) )
rethinking::drawdag( dag5.1 )
```

What this DAG says is:
(1) A directly influences D 
(2) M directly influences D 
(3) A directly influences M


## Interpreting DAGs: Testable implications of DAGs

$X \indep Y$


