---
title: "Chapter 5"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r include=FALSE}
# Setting Knit options for the document
knitr::opts_chunk$set( message=FALSE, warning=FALSE)
```
\newcommand{\indep}{\perp \!\!\! \perp}


# 5.1: Supurious Correlations
## We observe a puzzle: Marriage rates and divorce rates are positively correlated
```{r}
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce

d$W <- d$WaffleHouses
d$A <- scale(d$MedianAgeMarriage)
d$D <- scale(d$Divorce)
d$M <- scale(d$Marriage)
```

Let's fit a bayesian model to the three variables -- Waffle Houses, Age at marriage and Marriage Rate.

```{r}
waffle_divorce <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + bW*W,
    bW ~ dnorm(0, 0.2),
    a ~ dnorm(0, 0.2),
    sigma ~ dexp(1)
  ),
  data = d
)
age_divorce <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + bA*A,
    bA ~ dnorm(0, 0.2),
    a ~ dnorm(0, 0.2),
    sigma ~ dexp(1)
  ),
  data = d
)

marriage_divorce <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + bM*M,
    bM ~ dnorm(0, 0.2),
    a ~ dnorm(0, 0.2),
    sigma ~ dexp(1)
  ),
  data = d
)
```

```{r}
waffle_divorce_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(waffle_divorce) , Sigma=vcov(waffle_divorce))
waffle_seq <- seq(from=min(d$W)-1, to = max(d$W) + 1, length.out = 30)
input_matrix <- rbind(rep(1,length(waffle_seq)), waffle_seq)
mean_divorce_sim_m5_1 <- waffle_divorce_posterior[,c("a","bW")] %*% input_matrix
mean_divorce_PI_m5_1 <- apply(mean_divorce_sim_m5_1, 2, rethinking::PI, prob=0.89)

age_divorce_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(age_divorce) , Sigma=vcov(age_divorce))
age_seq <- seq(from=min(d$A)-1, to = max(d$A) + 1, length.out = 30)
input_matrix <- rbind(rep(1,length(age_seq)), age_seq)
mean_divorce_sim_m5_2 <- age_divorce_posterior[,c("a","bA")] %*% input_matrix
mean_divorce_PI_m5_2 <- apply(mean_divorce_sim_m5_2, 2, rethinking::PI, prob=0.89)

marriage_divorce_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(marriage_divorce) , Sigma=vcov(marriage_divorce))
marriage_seq <- seq(from=min(d$M)-1, to = max(d$M) + 1, length.out = 30)
input_matrix <- rbind(rep(1,length(marriage_seq)), marriage_seq)
mean_divorce_sim_m5_3 <- marriage_divorce_posterior[,c("a","bM")] %*% input_matrix
mean_divorce_PI_m5_3 <- apply(mean_divorce_sim_m5_3, 2, rethinking::PI, prob=0.89)

# Plot the graphs
par(mfcol=c(1,3))

plot(d$WaffleHouses, d$D, main="Waffle Houses and Divorce Rate")
abline(a=mean(waffle_divorce_posterior[,"a"]), b=mean(waffle_divorce_posterior[,"bW"]), lty=2)
shade(mean_divorce_PI_m5_1, waffle_seq)

plot(d$A, d$D, main="Age at marriage and Divorce Rate")
abline(a=mean(age_divorce_posterior[,"a"]), b=mean(age_divorce_posterior[,"bA"]), lty=2)
shade(mean_divorce_PI_m5_2, age_seq)

plot(d$M, d$D, main="Marriage Rate and Divorce Rate")
abline(a=mean(marriage_divorce_posterior[,"a"]), b=mean(marriage_divorce_posterior[,"bM"]), lty=2)
shade(mean_divorce_PI_m5_3, marriage_seq)
```


## Introduce some tools to solve the puzzle: DAGs
A DAG is a way of describing qualitative causal relationships among variables. It isn’t as detailed as a full model description, but it contains information that a purely statistical model does not. Unlike a statistical model, a DAG, if it is correct, will tell you the consequences of intervening to change a variable.

```{r}
library(dagitty)
dag5.1 <- dagitty( "dag {
    A -> D
    A -> M
    M -> D
}")
coordinates(dag5.1) <- list( x=c(A=0,D=1,M=2) , y=c(A=0,D=1,M=0) )

dag5.1_b <- dagitty( "dag {
    A -> D
    A -> M
}")
coordinates(dag5.1_b) <- list( x=c(A=0,D=1,M=1) , y=c(A=0,D=2,M=-2) )
par(mfcol=c(1,2))
rethinking::drawdag( dag5.1 )
rethinking::drawdag( dag5.1_b)
```

What the DAG on the left says is:
  (1) A directly influences D 
  (2) M directly influences D 
  (3) A directly influences M

The DAG on the right says:
  (1) A directly influences D 
  (2) M does not directly influence D 
  (3) A directly influences M


## Interpreting DAGs: Testable implications of DAGs

Comparing the DAG on the left to the DAG on the right, we can observe that in the right DAG, M influences D only due to the shared influence of A, i.e D and M share a common influence via A. 
If we condition for A, them M and D would be independent. A conclusion we represent as $D \indep M|A$

DAGs are analytical models of reality, not statistical models. This means that there are multiple potential statistical models that can fit the same DAG and it's implications. Here, we will use a linear model to condition for the effect of $A$ on $M$ and $D$. 

We test the two DAGs above using a multi-variate linear model

```{r}
multivar_model <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + bA*A + bM*M,
    a ~ dnorm(0, 0.2),
    bA ~ dnorm(0, 0.5),
    bM ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data=d
)

rethinking::precis(multivar_model)

# multivar_model_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(multivar_model) , Sigma=vcov(multivar_model))
# waffle_seq <- seq(from=min(d$W)-1, to = max(d$W) + 1, length.out = 30)
# input_matrix <- rbind(rep(1,length(waffle_seq)), waffle_seq)
# mean_divorce_sim_m5_1 <- waffle_divorce_posterior[,c("a","bW")] %*% input_matrix
# mean_divorce_PI_m5_1 <- apply(mean_divorce_sim_m5_1, 2, rethinking::PI, prob=0.89)
```
Note above that we do not simulate the posterior as usual. Why is this? 

This is because, multi-variate distributions are really difficult to simulate. Typical simulation involves generating a matrix of coeffcient values, and then multiplying that matrix by a predictor-value matrix to output the posteriod mean value for the outcome variable, ie

$$ MeanOutcomes = 

\begin{bmatrix}
\mathtt{Coeffienct Samples}
\end{bmatrix}

\begin{bmatrix}
\mathtt{PredictorValues}
\end{bmatrix} $$

The predictor-value matrix is easy to construct when there is only one variable. It's simply 

$$\begin{bmatrix}
\mathbf{1}, \mathtt{PredictorSequence}
\end{bmatrix}$$

where $\mathbf{1}$ is a vector with all 1's, and Predictor Sequence is a sequence of values drawn from the values of the predictor variable, typically via some code like `seq(from=min(Predictor), to=max(Predictor), by = 1)`.

However, when the no of predictors increase, the predictor value matrix becomes significantly larger. If we were to pick $n$ values for each predictor, then the predictor matrix would have $n*n$ rows. If we sample 1000 rows for the posterior coefficients matrix, and have k predictor variables, then our mean outcomes matrix will have 1000*$n^k$ rows. For n = 30 and k = 4, we are left with an 800 mn row matrix. 

## Understanding mulivariate predictions

As described above, generating the matrix of posterior predictions becomes harded as the number of predictors increases. The same applies to analysis which require summarising this matrix. 

Let's say we wish to plot the outcome variable against a predictor variable of interest $X_t$. In the single variate case, we would pick a sequence of $X_t$ values and generate posterior predictions for these values. However, now things have changed. To find the average $Y$ value at a particular $X_t$ value, we would need to account for the other $k-t$ predictors. We're back to the same problem as before, at each of the $n$ values of $X_t$ where we would like to evaluate the outcome, we need to manipulate a matrix with $1000*n^{k-1}$ rows. 

This does not mean that generating these outcomes is impossible. We can of course generating them using the vector of coefficients, and the Covaraince matrix. It's just that at this point in the text, we haven't developed the mathematical machinery to do this. As a result we need to look else where to generate predictive plots of outcomes. 

### Partialling out predictor variables

Consider a regression equation of the form $y = \alpha + \Sigma_i \beta_i X_i$ where $\beta_i$ denote the coefficients of the $i$ predictors. We'd like to understand the independent effect of one predictor, say $X_k$. How can we do this?

One way to do this is to partial out the variation in $X_k$ that can be explained by the other predictors. We do this by estimating the following regression: $Mean(X_k) = \alpha' + \Sigma_j\theta_jX_j$
Once we have this regression, the unexplained part of $X_k$ is $X_k - Mean(X_k)$, also known as the _residual_.

```{r}
m5.4 <- quap(
    alist(
        M ~ dnorm( mu , sigma ) ,
        mu <- a + bAM * A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bAM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
) , data = d )

m5.4_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(m5.4) , Sigma=vcov(m5.4))
# Note that instead of generating a sequence of predictor values, we're using the original data itself

m5.4_input_matrix <- t(cbind(rep(1,length(d$A)),d$A))
m5.4_Marriage_Rate_Predict <- m5.4_posterior[,c("a", "bAM")] %*% m5.4_input_matrix
m5.4_Marriage_Rate_mean <- apply(m5.4_Marriage_Rate_Predict, 2, mean)
m5.4_mean_Marriage_CI <- apply(m5.4_Marriage_Rate_Predict, 2, rethinking::PI, prob=0.89)
d$M_resid <- d$M - m5.4_Marriage_Rate_mean
plot(d$D ~ d$M_resid)
```

So, we've residualized the Marriage rate. We can now regress the Divorce rate on the residualized Marriage rate. 


```{r}
m5.4b <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM_resid * M_resid ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM_resid ~ dnorm( 0 , 0.25 ) ,
        sigma ~ dexp( 1 )
) , data = d )

m5.4b_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(m5.4b) , Sigma=vcov(m5.4b))

m5.4b_input_matrix <- t(cbind(rep(1,length(d$M_resid)), d$M_resid))
m5.4b_Divorce_Rate_Predict <- m5.4b_posterior[,c("a", "bM_resid")] %*% m5.4b_input_matrix
m5.4b_Divorce_Rate_mean <- apply(m5.4b_Divorce_Rate_Predict, 2, mean)
m5.4b_mean_Divorce_CI <- apply(m5.4b_Divorce_Rate_Predict, 2, rethinking::PI, prob=0.89)
plot(d$D ~ d$M_resid)
abline(a=mean(m5.4b_posterior[,"a"]),
       b=mean(m5.4b_posterior[,"bM_resid"]))
abline(v=0, lty=2)
shade(m5.4b_mean_Divorce_CI,d$M_resid)
```



### Goodness of Fit plots

The idea here is to plot predicted v/s actually observed values for outcomes to judge the fit. 

We start by fitting the model
```{r}
multivar_model <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + bA*A + bM*M,
    a ~ dnorm(0, 0.2),
    bA ~ dnorm(0, 0.5),
    bM ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data=d
)

# Extract a bunch of posterior parameter values
multivar_model_posterior <- MASS::mvrnorm( n=1e3 , mu=coef(multivar_model) , Sigma=vcov(multivar_model))
```

Due to the predictor value matrix becoming large, we'll stick to using the original data itself. 

```{r}
# Predictions = Posterior Parameters x Predictor Value matrix
predictor_val_matrix <- rbind(rep(1,length(d$A)), as.vector(d$A), as.vector(d$M))
mu_divorce_predictions <- multivar_model_posterior[,c("a", "bA", "bM")] %*% predictor_val_matrix
mu_divorce_mean <- apply(mu_divorce_predictions, 2, mean)
mu_divorce_PI <- apply(mu_divorce_predictions, 2, rethinking::PI, prob=0.89)
```

```{r}
plot( d$D, mu_divorce_mean, xlab="Observed Divorce", ylab="Predicted Divorce Rates")
abline( a=0 , b=1 , lty=2 )
for ( i in 1:nrow(d) ) lines( rep(d$D[i],2) , mu_divorce_PI[,i] , col=rangi2 ) # Drawss the 89% PI bars
```



